---
title: "How to Slippymath"
author: "Miles McBain"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

`slippymath` provides tools to assist your work with slippy map tile servers and the map tiles they provide. The main things it can help with are:
* determining what zoom level to use
* identifying the tiles you need to plot your data over
* identifying where lonlat coordinates fall in tile space
* determining the bounding boxes of tiles
* compositing tiles into a single spatially referenced raster.

Downloading tiles does not get a built-in helper. This is to make `slippymath`
as service agnostic as possible. There is an example of downloading from Mapbox
in the README. You may also enjoy the
[`ceramic`](https://github.com/hypertidy/ceramic) package by Michael Sumner.

## Determining zoom level

The main assumption `slippymath` makes is that you can calculate a bounding box
for your spatial data. Most spatial tools have this capability - see
`sf::st_bbox()`.

Given a bounding box in longitude and latitude (EPSG: 4326), you can execute a
'tile query' to see how many slippy map tiles your data would occupy at various
zoom levels:

```{r}
library(slippymath)

uluru_bbox <-
    c(xmin = 131.02084,
      xmax = 131.0535,
      ymin = -25.35461,
      ymax = -25.33568)

bb_tile_query(uluru_bbox)
```

When choosing your zoom level consider: Tiles are 256 x 256 pixel squares which
roughly range in size from 30 - 40kb. 100's of tiles will be 10's of megabytes.
You may also want to check rate limits imposed by the server you intend to pull
the tiles from.

`slippymath` can also choose a zoom level for your tiles, given a tile budget. See below.

## Identifying tiles to plot over

`slippymath` calls the set of tiles you need to plot all your data over a "Tile
Grid" or `tg` for short. You can get a tile grid given a bounding box and tile
zoom level OR a bounding box and tile budget (the zoom will be chosen to fit the
budget).

For example using the Uluru bounding box from above and zoom level 14:

```{r}
bb_to_tg(uluru_bbox, zoom = 14)
```

And using the same bounding box with a budget of 15 tiles:

```{r}
bb_to_tg(uluru_bbox, max_tiles = 15)
```

A tile grid is an object that contains a dataframe of tile coordinates (`$tiles`) and a `$zoom` level.

## Identifying the tile for a single point

Occasionally it may make sense to find the tile at a given zoom for a single
point. `slippymath` has functions to convert between longitude and latitude tile
coordinate space:

```{r}
lonlat_to_tilenum(131.02084, -25.35461, zoom = 14)

tilenum_to_lonlat(14154, 9385, zoom = 14)
```

In this example, the exact lon lat we started with is not returned in the
conversion from tile coordinate space. The lon lat returned is always the top
left corner point of the tile.

## Determining tile bounding boxes

Once you have tiles, you will likely want to know their spatial bounding boxes
as this is essential for compositing them. If you use the built-in compositing
function, this is taken care of for you.

Find the bounding boxes of a tile grid like so:

```{r}
uluru_grid <- bb_to_tg(uluru_bbox, max_tiles = 15)
tg_bbs(uluru_grid)

## for a single tile use tg_bb()
```

Slippy maps use a special mercator projection: EPSG: 3857. The tile bounding
boxes are expressed in this projection.

## Compositing tiles to a spatial raster

A simple function to composite tiles to a single raster with the correct spatial
extent (again in the EPSG 3857 projection) is provided. It constructs individual
raster and subsequently calls `raster::merge` which has issues with manual
calls to `gc` which can slow things down with many tiles.

To composite, it is assumed you have a tile grid object, and a list of image
files that correspond in order to the tiles in the tile grid. See the README
example for a way to download tiles and get such a list.

```{r, eval=FALSE}
raster_out <- tg_composite(tile_grid, images)
```

Finally a convenience function is provided to convert the raster to a .png if needed:

```{r, eval=FALSE}
raster_to_png(raster_out, "uluru.png")
```
